Spanner: Google's Globally-Distributed Database \newline Seminar Report

### 2016-02-25 paper/paper.tex
Iterate section 4

The TrueTime API allows Spanner to guarantee some correctness properties of the
implemented concurrency control. These enable some of the main features, i.e.,
externally-consistent transactions, lock-free read-only transactions and so-called
\emph{snapshot reads}. Snapshot reads are non-blocking reads in the past at a
specific timestamp $t$. In this case, Spanner guarantees that snapshot reads will
see the effects of all transactions committed as of $t$.

Spanner provides different operations, as can be seen in
Table~\ref{tbl:reads-writes}. Read-only transactions have to be made explicit.
This requires the application/user to predeclare that the read-only transaction
does not include write operations. Moreover, read-only transaction have the
benefits of snapshot isolation because they are \emph{lock-free}. Standalone
writes are simply read-write transaction without any read operations. Non-snapshot
standalone reads are implemented as read-only transactions. Read-only transactions
and snapshot reads proceed on any \emph{sufficiently up-to-date} replica. For
snapshot reads, the application/user has two options. On the one hand, a precise
timestamp can be specified at which the read should be executed. On the other
hand, an upper bound can be specified and Spanner choses a timestamp that
satisfies this bound.

\begin{table}
  \centering
  \begin{tabular}{|l||l|l|}
    \hline
    {\bfseries Operation} & {\bfseries Concurrency Control} &
    {\bfseries Replica required} \tabularnewline
    \hline\hline
    Read-Write transaction & Pessimistic & leader \tabularnewline
    \hline
    Read-Only transaction & Lock-free & leader; any
    \tabularnewline
    \hline
    Snapshot Read, client-provided timestamp & Lock-free & any \tabularnewline
    \hline
    Snapshot Read, client-provided bound & Lock-free & any \tabularnewline
    \hline
  \end{tabular}
  \caption{Reads and writes in Spanner.}
  \label{tbl:reads-writes}
\end{table}

@@

Read-write transactions use two-phase locking and Spanner may only assign
timestamps in between the two phases. Spanner enforces another invariant, i.e.,
a \emph{monotonicity invariant}, which makes use of the disjointness invariant.
This monotonicity invariant guarantees that within each Paxos group, Spanner
assigns timestamps to Paxos writes in monotonically increasing order. To guarantee
this across leaders, Spanner utilizes the disjointness invariant. This implies
that a leader must only assign timestamps within the interval of its leader
lease. $s_{max}$ is advanced to $s$ whenever a timestamp $s$ is assigned by a
leader.

To guarantee externally-consistent reads and writes, Spanner enforces an
\emph{external consistency invariant}. In essence, this invariant guarantees that
whenever the commit of a transaction $T_1$, denoted as $e_1^{commit}$, occurs
before the start of a transaction $T_2$, denoted as $e_2^{start}$, then the commit
timestamp of $T_1$, denoted as $s_1$, is guaranteed to be smaller than the
timestamp of $T_2$, denoted as $s_2$. More formally, the following must hold to
guarantee external consistency:
\begin{center}
$t_{abs}\left(e_1^{commit}\right) < t_{abs}\left(e_2^{start}\right) \Rightarrow s_1 < s_2$
\qquad\qquad $t_{abs}(e)$ \ldots absolute time of an event $e$
\end{center}

\noindent
Spanner obeys two rules in order to guarantee this invariant:
\begin{description}
  \item[Start] Coordinator leader for a write $T_i$ assigns a commit timestamp
    $s_i \geq TT.now().latest$, where $TT.now()$ is executed right after the
    arrival event of the commit request for $T_i$.
  \item[Commit Wait] Coordinator leader ensures that clients cannot see data
    committed by $T_i$ until $TT.after\left(s_i\right)$ is true. This ensures
    that the assigned timestamp $s_i$ is less than the absolute commit time of
    $T_i$.~\cite{Corbett:2012}
\end{description}

>>>>

The TrueTime API allows Spanner to guarantee some correctness properties of the
implemented concurrency control. These enable some of the main features:
externally-consistent transactions, lock-free read-only transactions, and
so-called \emph{snapshot reads}. Snapshot reads are non-blocking reads in the
past at a specific timestamp $t$. Spanner guarantees that snapshot reads will
see the effects of all transactions committed as of $t$.

Spanner provides different operations, as shown in Table~\ref{tbl:reads-writes}.
Read-only transactions have to be made explicit. This requires the
application/user to predeclare that the read-only transaction does not include
write operations. Moreover, read-only transaction have the benefits of snapshot
isolation because they are \emph{lock-free}. Standalone writes are simply
read-write transaction without any read operations. Non-snapshot standalone reads
are implemented as read-only transactions. In contrast to snapshot reads, which
may be executed on any timestamp in the past, read-only transactions execute on
the last committed timestamp. Both, snapshot reads and read-only transactions,
proceed on any \emph{sufficiently up-to-date} replica. For snapshot reads,
the application/user has two options. On the one hand, a precise timestamp can
be specified at which the read should be executed. On the other hand, an upper
bound can be specified, and Spanner choses a timestamp that satisfies this bound.

\begin{table}[ht]
  \centering
  \begin{tabular}{|l||l|l|}
    \hline
    {\bfseries Operation} & {\bfseries Concurrency Control} &
    {\bfseries Replica required} \tabularnewline
    \hline\hline
    Read-Write transaction & Pessimistic & leader \tabularnewline
    \hline
    Read-Only transaction & Lock-free & leader; any
    \tabularnewline
    \hline
    Snapshot Read, client-provided timestamp & Lock-free & any \tabularnewline
    \hline
    Snapshot Read, client-provided bound & Lock-free & any \tabularnewline
    \hline
  \end{tabular}
  \caption{Reads and writes in Spanner.}
  \label{tbl:reads-writes}
\end{table}

@@

Read-write transactions use two-phase locking, and Spanner may only assign
timestamps in between the two phases. Spanner enforces another invariant, the
a \emph{monotonicity invariant}, which makes use of the disjointness invariant.
This monotonicity invariant guarantees that within each Paxos group, Spanner
assigns timestamps to Paxos writes in monotonically increasing order. To guarantee
this across leaders, Spanner utilizes the disjointness invariant. This implies
that a leader must only assign timestamps within the interval of its leader
lease. $s_{max}$ is advanced to $s$ whenever a timestamp $s$ is assigned by a
leader.

To guarantee externally-consistent reads and writes, Spanner enforces an
\emph{external consistency invariant}. In essence, this invariant guarantees that
whenever the commit of a transaction $T_1$, denoted as $e_1^{commit}$, in real
time occurs before the start of a transaction $T_2$, denoted as $e_2^{start}$,
then the commit timestamp of $T_1$, denoted as $s_1$, is guaranteed to be smaller
than the commit timestamp of $T_2$, denoted as $s_2$. More formally, the following
must hold to guarantee external consistency:
\begin{center}
$t_{abs}\left(e_1^{commit}\right) < t_{abs}\left(e_2^{start}\right) \Rightarrow s_1 < s_2$
\qquad\qquad $t_{abs}(e)$ \ldots absolute time of an event $e$
\end{center}

\noindent
Spanner obeys two rules in order to guarantee this invariant:
\begin{description}
  \item[Start] Coordinator leader for a write transaction $T_i$ assigns a commit
    timestamp $s_i \geq TT.now().latest$, where $TT.now()$ is executed right after
    the arrival event of the commit request for $T_i$.
  \item[Commit Wait] Coordinator leader ensures that clients cannot see data
    committed by $T_i$ until $TT.after\left(s_i\right)$ is true. This ensures
    that the assigned timestamp $s_i$ is less than the absolute commit time of
    $T_i$.~\cite{Corbett:2012}
\end{description}
